Data Specialization for a Staged Lambda Calculus

### 2015-02-20 paper/example.tex

\crem{
Notes from 1/16:
\begin{itemize}
\item We are modeling computations where some inputs arrive at stage 1 and
others arrive at stage 2. As a result, some subcomputations can be performed at
stage 1, and others at stage 2.
\item \emph{Statics}: M12 is a language with three worlds: 1M, 1P, and 2.
We model the computations described above as functions in world M, and their
results as values in 2.
\item \emph{Dynamics}: We describe how to perform all of the stage 1
subcomputations in an M12 program. Our semantics are mutually defined in terms
of how to reduce terms in each of the three worlds.
\end{itemize}
}

\crem{Now that hold is not officially in the language, I think we have to put a
function prototype in the quickselect code figure. (And move the footnote about
it to the end of 3.1.}

@@

The $\next$ wraps a second-stage expression of type $\rmint$ (the \texttt{case}
expression) as a first-stage expression of type $\fut\rmint$ (the result type
of \texttt{qsStaged}). Inside the \texttt{LT} and \texttt{GT} branches, the
$\prev$ unwraps first-stage $\fut\rmint$s (the results of the recursive calls
to \texttt{qsStaged}) as second-stage $\rmint$s. In the \texttt{EQ} branch,
$\pause$ promotes a first-stage $\rmint$ (the head of \texttt{l}) directly to a
second-stage $\rmint$.%
\footnote{$\pause$ can be implemented from our other language features, with
some effort. We provide it as a primitive to shorten examples. Furthermore, it
would be wise to extend $\pause$ to other base types, if we had them, and to
products and sums thereof.  This is related to the notion of {\em mobility} in
\cite{murphy05} and {\em stability} in \cite{krishnaswami13}.}

The process of adding staging annotations ($\fut$ types, $\next$, $\prev$, and
$\pause$) to unstaged code has been the subject of extensive research under the
name of \emph{binding time analysis}. In this paper, we assume that these
annotations have been provided by the programmer (or perhaps a binding time
analysis tool), and do not consider the problem of generating such annotations.
There are many ways to annotate any program, including \texttt{qSelect}; we
chose annotations which maximize the work performed in the first stage.

>>>>

@@

The $\next$ wraps a second-stage expression of type $\rmint$ (the \texttt{case}
expression) as a first-stage expression of type $\fut\rmint$ (the result 
of \texttt{qsStaged}). Inside the \texttt{LT} and \texttt{GT} branches, the
$\prev$ unwraps first-stage $\fut\rmint$s (the results of the recursive calls
to \texttt{qsStaged}) as second-stage $\rmint$s. In the \texttt{EQ} branch,
$\pause$ promotes a first-stage $\rmint$ (the head of \texttt{l}) directly to a
second-stage $\rmint$.

The process of adding staging annotations ($\fut$ types, $\next$, and $\prev$)
to unstaged code has been the subject of extensive research under the
name of \emph{binding time analysis}. In this paper, we assume that these
annotations have been provided by the programmer (or perhaps a binding time
analysis tool), and do not consider the problem of generating such annotations.
There are many ways to annotate any program, including \texttt{qSelect}; we
chose annotations which maximize the work performed in the first stage.

### 2015-02-20 paper/semantics.tex

The example code in this paper makes liberal use of $\rmint$s and various
functions on these, as well as a function $\pause$ which takes a $\curr\rmint$
to a $\fut\rmint$.

>>>>

The example code in this paper makes liberal use of $\rmint$s and various
functions on these, as well as a function $\pause$ which takes a $\curr\rmint$
to a $\fut\rmint$.%
\footnote{$\pause$ is definable in \lang\ given an inductive definition of
$\rmint$s. In practice, we provide both $\rmint$s and $\pause$ as primitives.
It is sensible to extend $\pause$ to all base types and to products and sums
thereof. This is related to the notion of {\em mobility} in \cite{murphy05} and
{\em stability} in \cite{krishnaswami13}.}
